---
title: "Group_SARIMAX"
author: "Group_1_4"
date: "12/1/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tseries)
library(forecast)
library(tidyverse)
library(vars)
```

## Load data and visualize data   
   
```{r}
setwd('~/classes/TimeSeries/group_project')
train = read.csv('train.csv')
test = read.csv('test.csv')

# Split train data to train and validataion data, roughly based on 8:2 rule  
train_tr = train[1:288,]
train_tv = train[289:336,]

## train
BankrupRate_tr <- ts(train_tr$Bankruptcy_Rate, 
                   start = c(1987,1), end = c(2010,12), frequency = 12)
UnempRate_tr <- ts(train_tr$Unemployment_Rate, 
                   start = c(1987,1), end = c(2010,12), frequency = 12)
Pop_tr <- ts(train_tr$Population, 
                   start = c(1987,1), end = c(2010,12), frequency = 12)
HousePriIdx_tr <- ts(train_tr$House_Price_Index, 
                   start = c(1987,1), end = c(2010,12), frequency = 12)
# validation
BankrupRate_tv <- ts(train_tv$Bankruptcy_Rate, 
                   start = c(2011,1), end = c(2014,12), frequency = 12)
UnempRate_tv <- ts(train_tv$Unemployment_Rate, 
                   start = c(2011,1), end = c(2014,12), frequency = 12)
Pop_tv <- ts(train_tv$Population, 
                   start = c(2011,1), end = c(2014,12), frequency = 12)
HousePriIdx_tv <- ts(train_tv$House_Price_Index, 
                   start = c(2011,1), end = c(2014,12), frequency = 12)

# test
UnempRate_te <- ts(test$Unemployment_Rate, 
                   start = c(2015,1), end = c(2017,12), frequency = 12)
Pop_te <- ts(test$Population, 
                   start = c(2015,1), end = c(2017,12), frequency = 12)
HousePriIdx_te <- ts(test$House_Price_Index, 
                   start = c(2015,1), end = c(2017,12), frequency = 12)

# Train
BankrupRate_t <- ts(train$Bankruptcy_Rate, 
                   start = c(1987,1), end = c(2014,12), frequency = 12)
UnempRate_t <- ts(train$Unemployment_Rate, 
                   start = c(1987,1), end = c(2014,12), frequency = 12)
Pop_t <- ts(train$Population, 
                   start = c(1987,1), end = c(2014,12), frequency = 12)
HousePriIdx_t <- ts(train$House_Price_Index, 
                   start = c(1987,1), end = c(2014,12), frequency = 12)
```
  
```{r}
plot(ts(train$Bankruptcy_Rate, 
                   start = c(1987,1), end = c(2014,12), frequency = 12),
     ylab="Bankruptcy Rate (%)", xlab="years")
par(mfrow=c(3,1))
plot(ts(train$Unemployment_Rate, 
                   start = c(1987,1), end = c(2014,12), frequency = 12),
     ylab="Unemployment Rate (%)", xlab="years")
plot(ts(train$House_Price_Index, 
                   start = c(1987,1), end = c(2014,12), frequency = 12),
     ylab="House Price Index", xlab="years")
plot(ts(train$Population, 
                   start = c(1987,1), end = c(2014,12), frequency = 12),
     ylab="Population", xlab="years")

acf(ts(train$Bankruptcy_Rate, 
                   start = c(1987,1), end = c(2014,12), frequency = 12))

```
  
    
```{r}
# Visualize each time series
par(mfrow=c(2,2))
plot(BankrupRate_tr)
plot(UnempRate_tr)
plot(Pop_tr)
plot(HousePriIdx_tr)
```
    
Bankruptcy rate is the variable we concerned, having both seasonality and trend.  
      
```{r}
par(mfrow=c(2,2))
acf(BankrupRate_tr, lag.max = 72)
ccf(BankrupRate_tr, UnempRate_tr, lag.max = 60)
ccf(BankrupRate_tr, Pop_tr, lag.max = 60)
ccf(BankrupRate_tr, HousePriIdx_tr, lag.max = 60)
```
      
From the acf plot, the bankruptcy rate does show the trend and seasonality. According to the ccf plots, all the variables are correlated to bankfruptcy rate.      
          
## Fit Models and evaluate it based on rmse (Should include breif reasoning and graphs to support, also model diagonal if applicable)    
     
### select d, m and D       
```{r}
# Transformation
plot(BankrupRate_tr) # heteroscedasticity
(blambda <- BoxCox.lambda(BankrupRate_tr)) # 0.2055198
trans_BankrupRate_tr <- BoxCox(BankrupRate_tr, lambda = blambda)
qqnorm(trans_BankrupRate_tr);qqline(trans_BankrupRate_tr, col="red")
# d=1
ndiffs(trans_BankrupRate_tr) # 1
d1_BankrupRate_tr <- diff(trans_BankrupRate_tr)
adf.test(d1_BankrupRate_tr)
plot(d1_BankrupRate_tr) 
acf(d1_BankrupRate_tr, lag.max = 72) # lag12
# D=1 or 2
d1.12_BankrupRate <- diff(d1_BankrupRate_tr, lag = 12)
plot(d1.12_BankrupRate)
acf(d1.12_BankrupRate, lag=120)
d1.12.2_BankrupRate <- diff(d1.12_BankrupRate, lag = 12)
plot(d1.12.2_BankrupRate)
acf(d1.12.2_BankrupRate, lag=120)
```
      
### Select order p, q, P, Q   
    
```{r}
acf(d1.12_BankrupRate, lag.max = 72)
acf(d1.12.2_BankrupRate, lag.max = 96) # q=1,5, Q=2,4,5,6
pacf(d1.12.2_BankrupRate, lag.max = 72) # p=2, 3, 4 P=1, 2, 4
```

### fit models    
```{r, message=FALSE, warning=FALSE}
# q=1,5, Q=1,2,4
# p=2, 3, 4 P=1,2,4
# d = 1, D=1,2, period=12
#we should fit CSS model because of heterogeneous
i <-  1
rmse <- 1:(3*3*3*4*2)
models <- list()
for (p in c(2, 3, 4)) {
  for (P in c(1, 2, 4)) {
    for(q in c(1, 3, 5)) {
      for(Q in c(2, 4, 5, 6)) {
        for (D in c(1, 2)) {
        tryCatch({model <- Arima(BankrupRate_tr, order = c(p,1,q),
                       seasonal = list(order = c(P, D, Q), period=12),
                       xreg = data.frame(
                         UnempRate=UnempRate_tr, 
                         HousePriIdx = HousePriIdx_tr, 
                         Pop = Pop_tr), 
                       lambda = blambda,
                       method = "CSS")},
                 error = function(model) {
                   print(model)
                 })
        models[i] <- model
        y <- forecast(model, h=48, xreg = data.frame(
          UnempRate=UnempRate_tv, 
          HousePriIdx = HousePriIdx_tv, 
          Pop = Pop_tv))
        rmse[i] <- sqrt(mean((y$mean - BankrupRate_tv)^2))
        i <- i + 1
        }
      }
    }
  }
}

#create name
i = 1
models_name <- list()
for (p in c(2, 3, 4)) {
  for (P in c(1, 2, 4)) {
    for(q in c(1, 3, 5)) {
      for(Q in c(2, 4, 5, 6)) {
        for (D in c(1, 2)) {
        order <- paste(p,'1', q, sep="")
        season <- paste(P, D, Q, sep="")
        name <- paste(order, season, sep="_")
        models_name[i] <- name
        i <- i + 1
        }
      }
    }
  }
}
models_name = unlist(models_name)
df <- data.frame(order = models_name, rmse=rmse)
df[which(df$rmse < 0.24),] # best one is order 415_212
```
    
### best model diagonal and visualization
```{r}
# 415_212
best_model <- Arima(BankrupRate_tr, order = c(4,1,5),
                       seasonal = list(order = c(2, 1, 2), period=12),
                       xreg = data.frame(
                         UnempRate=UnempRate_tr, 
                         HousePriIdx = HousePriIdx_tr, 
                         Pop = Pop_tr), 
                       lambda = blambda,
                       method = "CSS")

y <- forecast(best_model, h=48, xreg = data.frame(
          UnempRate=UnempRate_tv, 
          HousePriIdx = HousePriIdx_tv,
          Pop = Pop_tv))

rmse_best <- sqrt(mean((y$mean - BankrupRate_tv)^2))
  
tsdiag(best_model) # assumptions semms reasonable

f <- BankrupRate_tr - best_model$residuals
{
  plot(BankrupRate_tr, type = "l", main = "SARIMAX Model for Bankruptcy Rate", 
      ylab = "Percentage Point", xlab = "Year",lwd=2)
  points(f, type = "l", col = "red")
  legend("bottomright", legend = c("Observed", "Predicted"), 
         lty = 1, lwd = 2:1, col = c("black", "red"), cex = 0.5)}

y_hat <- forecast(best_model, h=48, xreg = data.frame(
          UnempRate=UnempRate_tv, 
          HousePriIdx = HousePriIdx_tv, 
          Pop = Pop_tv))
{
  plot(y_hat)
  points(BankrupRate_tv, type = "l", col = "red")
  legend("bottomright", legend = c("predicted", "observed"), 
         lty = 1, col = c("blue", "red"), cex = 0.5)}
  
```
               
```{r}
par(mfrow=c(2,1))
{plot(best_model$residuals, main = "Residuals vs. Time", ylab = "Residuals")
abline(h = 0, col = "red")
acf(best_model$residuals, main = "ACF of Residuals")}
```
               
                  
```{r}
best_model_entire <- Arima(BankrupRate_t, order = c(4,1,5),
                       seasonal = list(order = c(2, 1, 2), period=12),
                       xreg = data.frame(
                         UnempRate=UnempRate_t, 
                         HousePriIdx = HousePriIdx_t, 
                         Pop = Pop_t), 
                       lambda = blambda,
                       method = "CSS")

y_hat_entire <- forecast(best_model_entire, h=36, xreg = data.frame(
          UnempRate=UnempRate_te, 
          HousePriIdx = HousePriIdx_te, 
          Pop = Pop_te))

plot(y_hat_entire, main = "Forecast from 2015-2017 with SARIMAX(4,1,5)*(2,1,2)[12] model")

predict_final <- as.numeric(y_hat_entire$mean)
predict_final <- round(predict_final, digits=4)

write.table(predict_final, "Team1.4forecasts.txt", sep="\n", row.names=F, col.names = F)
```
   
        

    
    


       

